---
title: "Pacific Salmon SAR & SAS Estimator"
author: "Kris Warner & Matt Siskey"
format: 
  html:
    toc: true
    theme: sketchty
    embed-resources: true
    self-contained-math: true
    code-tools: true
    code-fold: true
editor: visual
echo: true
params:
  brood_yr: 2024
  sp: "1" #"Coho" "Chum" "Pink"
  hatchery: "3F42001  270168 H"  #	3F42001  260002 H02; 3F42001  270168 H
  yr_trim: true
  yr_window: !expr c(seq(2013,2025,1)) 
---

## Setup

This script uses WDFW hatchery release data and hatchery rack return data to fit program-specific return-per-release models for Pacific salmon. Release and return data are categorized by release size, release stage, release location and uniquely idetified by coded-wire-tag codes. Return-per-release time series are then fit to generalized additive models to summarize and identify potential significant covariates. The first chunk below loads libraries, sets directories, and sources any functions required for the analysis.

```{r setup}
##############################
### Clear workspace & load libraries
# rm(list=ls())
# # First-time rmisr install
# options(download.file.method = "wininet"); library(devtools); devtools::install_github("MattCallahan-NOAA/rmisr", force = TRUE)

# Load packages, several aren't used in this workflow
pacman::p_load(rmisr, dplyr, purrr, stringr, tidyverse, knitr, kableExtra, ggplot2, RColorBrewer, 
               jsonlite, rmisr, janitor, keyring, mgcv, here, lubridate, tidygam, FSSgam)
# detach("package:plyr", unload = TRUE)

# Supply API key through keyring to access RMIS server & maintain credential security
token <- key_get("rmis_api") # Kris had a "token" in key_get too -- whatfer?
# set plotting theme
theme_set(theme_minimal())

##############################
### Set up drives & controls
my_drive    <-paste0("C:/Users/sism1477/OneDrive - Washington State Executive Branch Agencies/")
share_drive <-paste0(my_drive,"DFW-Team FP Westside Hatchery MandE - R/")
proj_drive  <-paste0(share_drive,"psalmon_sar_sas/")
`%nin%`     <-negate(`%in%`) # create an exclude operator
source(paste0(proj_drive,"functions/functions.R",sep="")) #Source functions for use below

```

## Pull data, and format

This chunk pulls location, release, and recovery data from the RMIS server for the species and hatchery stock of interest. Smolt-adult return (SAR) is calculated as the proportion of released fish that return to a designated location (e.g., the hatchery rack). The brood year release will not all return to the rack, and those fish that do return to the rack will not all return the same year. Once the window for a brood year's return has finished (6? years post-release), the SAR can be calculated. However, if the brood year release is not 100% marked, an expansion must be done on the recovery data based on the estimated as the summation of tagged fish released expanded by the hatchery program's mark rate (?).

Notes:

-   Data filtering

    -   Release data used excludes untagged release groups (i.e., tag_code_or_release_id begins with "!")
    -   Recovery data only include CWT-tagged fish
    -   Survival is calculated for tagged fish, then survival rate is applied to total release numbers to get the total return

-   releases: event_released \<-cwt_1st_mark_count + cwt_2nd_mark_count

-   returns: cwt_recovered \<-aggegrate(number_cwt_estimated \~ brood_year + tag_code, data = rec_dat, FUN = sum)

```{r data, results='hide', echo=FALSE, message=FALSE, warning=FALSE}
##############################
# ### Generate data for model fitting
# # retrieve all location data from rmis
# loc_all <- get_location(token = token)
# 
# #split location data into LUTS by location type. This could be a future application of a helper function
# # Location Type 4 = release site, Location Type 3 = Hatchery/Facility; Location Type 1 = Recovery Site
# 
# #release LUT
# relt <- loc_all |>
#   filter(location_type == 4) |>
#   # rename(release_location_name = name,
#   #        release_location_description = description,
#   #        release_location_code = location_code) |>
#   select(-c(id, format_version, record_code))
# 
# ##recovery LUT
# rect <- loc_all |>
#   filter(location_type == 1) |>
#   # rename(recovery_location_name = name,
#   #        recovery_description = description,
#   #        recovery_location_code = location_code) |>
#   select(-c(id, format_version, record_code))
# 
# #hatchery LUT
# hatt <- loc_all |>
#   filter(location_type == 3) |>
#   # rename(hatchery_name = name,
#   #        hatchery_description = description,
#   #        hatchery_location_code = location_code) |>
#   select(-c(id, format_version, record_code))
# 
# #stock LUT
# stockt <- loc_all |>
#   filter(location_type == 5) |>
#   # rename(stock_name = name,
#   #        stock_location_code = location_code,
#   #        stock_description = description) |>
#   select(-c(id, format_version, record_code))


##############################
### [RELEASES] Pull in RMIS release data
# rel_wide <-get_releases(drive, sy_stages, yrl_stages, smolt_stages)

### hatchery of interest location codes, found on the hatt data frame
# 180k+ release records total
rel_dat <-rmisr::get_release(token = token,
                      # brood_year = params$brood_yr,
                      species = params$sp,
                      hatchery_location_code = params$hatchery)

rel_dat[is.na(rel_dat$release_stage),"release_stage"] <- "no_stage"

rel_dat[is.na(rel_dat$cwt_1st_mark_count),"cwt_1st_mark_count"] <-0
rel_dat$cwt_1st_mark_count <-as.numeric(rel_dat$cwt_1st_mark_count)
rel_dat[is.na(rel_dat$cwt_2nd_mark_count),"cwt_2nd_mark_count"] <-0
rel_dat$cwt_2nd_mark_count <-as.numeric(rel_dat$cwt_2nd_mark_count)
rel_dat[is.na(rel_dat$non_cwt_1st_mark_count),"non_cwt_1st_mark_count"] <-0
rel_dat$non_cwt_1st_mark_count <-as.numeric(rel_dat$non_cwt_1st_mark_count)
rel_dat[is.na(rel_dat$non_cwt_2nd_mark_count),"non_cwt_2nd_mark_count"] <-0
rel_dat$non_cwt_2nd_mark_count <-as.numeric(rel_dat$non_cwt_2nd_mark_count)

rel_dat$event_released <-rel_dat$cwt_1st_mark_count +
                         rel_dat$cwt_2nd_mark_count #+
                         # rel_dat$non_cwt_1st_mark_count +
                         # rel_dat$non_cwt_2nd_mark_count

rel_dat$release_stage <-as.factor(rel_dat$release_stage)


rel_agg <-rel_dat %>% filter(!grepl("!",tag_code_or_release_id)) %>%
                      mutate(first_release_date = make_date(first_release_date_year, first_release_date_month, first_release_date_day)) %>%
                      mutate(jday = yday(first_release_date)) %>%
                      aggregate(event_released~species+brood_year+release_location_code+hatchery_location_code+tag_code_or_release_id+
                                               first_release_date_year+first_release_date_month+first_release_date_day+
                                               first_release_date+jday+avg_weight+release_stage, 
                                data=., FUN=sum) %>%
                      rename(tag_code = tag_code_or_release_id)

if(params$yr_trim=="TRUE"){rel_agg <-filter(rel_agg, brood_year %in% params$yr_window)}

##############################
# [RETURNS] Pull in WDFW FishBooks return data
# returns_HOR_agg <-get_returns(drive, wria_region)

# use the tag codes from the release query to grab recoveries; pull capped at 500k records
rel_agg_cwt <-unique(rel_agg$tag_code)
rec_cwt <-list()

for(t in 1:length(rel_agg_cwt)){
  rec_cwt[[t]] <-get_recovery(token = token, tag_code = rel_agg_cwt[[t]], species = params$sp)
}

names(rec_cwt) <-rel_agg_cwt
rec_cwt_df <-map_df(rec_cwt, ~as.data.frame(.))

rec_agg <-rec_cwt_df %>%  mutate(recovery_date = make_date(recovery_date_year, recovery_date_month, recovery_date_day)) %>%
                          mutate(jday = yday(recovery_date)) %>%
                          aggregate(number_cwt_estimated~species+run_year+recovery_date+jday+recovery_location_code+fishery+tag_code,
                                     data=., FUN=sum) %>%
                          filter(., recovery_location_code == params$hatchery)

rec_join <-aggregate(number_cwt_estimated~species+tag_code, data=rec_agg, FUN=sum)




# #################################################################################
# ######################### Missing expansion corrections #########################
# # See how many recoveries occured across 'fishery' categories
# table(rec_dat_agg$fishery)
# 
# ### RMIS 'fishery' types
# # 10-19 Troll
# # 20-29 Net and Seine
# # 30-39 Aboriginal
# # 40-49 Sport
# # 50-59 Escapement
# ## 50 Hatchery rack
# ## 52 Fish trap
# ## 54 Spawning ground
# # 60-69 Test Fisheries
# # 70-79 Juvenile Sampling
# # 80-89 High Seas
# # 90-99 Miscellaneous
# 
# ##############################
# ### Correct the instances of number_cwt_estimated == c(0,NA)
# prob_child <-filter(rec_dat_agg, number_cwt_estimated == 0 | number_cwt_estimated == "NA")
# good_child <-filter(rec_dat_agg, number_cwt_estimated != 0)
# # nrow(rec_dat_agg);nrow(good_child);nrow(prob_child)
# 
# # Fishery harvest:
# # (1)
# rec_dat_fishery <-filter(prob_child, fishery %in% c(seq(10,49,1),seq(80,89,1)))
# nrow(rec_dat_fishery)
# 
# # Hatchery rack
# rec_dat_rack <-filter(prob_child, fishery == 50)
# nrow(rec_dat_rack)
# 
# # Spawning ground
# rec_dat_sgs <-filter(prob_child, fishery == 54)
# nrow(rec_dat_sgs)
# 
# # fishery-independent survey & misc
# rec_dat_misc <-filter(prob_child, fishery %in% c(51,52,53,seq(55,79,1),seq(90,99,1)))
# nrow(rec_dat_misc)
#
##############################
# CatchSample (exp)
# rec_dat$catch_sample_id ==NA -> number_cwt_estimated==0
## 1 fish expands to 0 fish
# page 80 of rmis data spec explains expansion
# cs <-get_catchsample(token=token,
#                      catch_year=2011,
#                      species=2)
#
# #################################################################################



########################## Join Release & Return Data for SAR ##########################
sar_df <-left_join(rel_agg, rec_join[,c("tag_code","number_cwt_estimated")], by = "tag_code") %>%
         mutate(avg_weight = round(avg_weight, digits = -1)) %>%
         aggregate(cbind(event_released,number_cwt_estimated)~species+brood_year+release_location_code+hatchery_location_code+
                                                              first_release_date_month+jday+avg_weight+release_stage,
                   data=., FUN=sum)

sar_df$sar <-sar_df$number_cwt_estimated/sar_df$event_released
# sar_df <-sar_df[!is.na(sar_df$sar),]


####################################
### sar by brood year
ggplot()+
    geom_boxplot(data=sar_df,aes(x=brood_year,y=sar, group=brood_year))+
    scale_x_continuous(breaks=seq(min(sar_df$brood_year),max(sar_df$brood_year),1))

### sar by release stage
ggplot()+
    geom_boxplot(data=sar_df,aes(x=release_stage,y=sar, group=release_stage))

### sar by release size (avg_weight)
ggplot(data=sar_df,aes(x=avg_weight,y=sar))+
    geom_boxplot(data=sar_df,aes(x=avg_weight,y=sar, group=avg_weight))

### sar by release timing (jday)
ggplot(data=sar_df,aes(x=jday,y=sar))+
    geom_bar(stat="identity")

ggplot(data=sar_df,aes(x=as.factor(first_release_date_month),y=sar))+
    geom_boxplot(data=sar_df,aes(x=as.factor(first_release_date_month),y=sar, group=as.factor(first_release_date_month)))

### sar by release location
ggplot(data=sar_df,aes(x=release_location_code,y=sar))+
    geom_boxplot(data=sar_df,aes(x=release_location_code,y=sar, group=release_location_code))


```

## Fit SAR Models

There are a variety of methods that can be employed for summarizing and forecasting SAR. A simple form of this would entail fitting a generalized additive model to the time series of SAR data. Here, we use a different form of this and fit quasi-binomial generalized additive models to the time series of two components of the smolt-adult-return data: total tags released & total tags lost. Return-per-release (or SAR) is explored with the following covariates in mind:

1.  Total return-per-release by brood year
2.  Release stage
3.  Release timing
4.  Release size
5.  Release location

```{r fit}
################################################################################
##### Generate model sets 
# # factor.vars = c()
# cont.vars = c("brood_year", "avg_weight", "jday", "first_release_date_month") #"first_release_date_month" "avg_weight"
# 
# max.knots <--1
# max.preds <-4
# max.fits  <-2000
# 
# 
# start.fit = gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ brood_year,
#                 family=quasibinomial(link = "logit"),
#                 data=sar_df,
#                 method="REML")
# 
# 
# model.set = generate.model.set(use.dat                    = sar_df,
#                                test.fit                   = start.fit,
#                                pred.vars.cont             = cont.vars,
#                                # pred.vars.fact             = factor.vars,
#                                k                          = max.knots, #Select df or maximum number of knots
#                                factor.factor.interactions = T, #Remove interactions
#                                smooth.smooth.interactions = T, #Remove interactions
#                                max.predictors             = max.preds) #Select Maximum Number of Predictors
# 
# out.list = fit.model.set(model.set, parallel=T, max.models = max.fits)
# 
# # look at the model selection table
# mod.table = out.list$mod.data.out
# mod.table = mod.table[order(mod.table$AICc),]
# best.model.aic = out.list$success.models[[as.character(mod.table$modname[1])]]
# second.model.aic = out.list$success.models[[as.character(mod.table$modname[2])]]


################################################################################
# [FIT_DAT] Prepare datasets for model fitting
# fit_dat <-make_fit_dat(rel_wide,age_comps,age_comps_w,returns_HOR_agg)

#----------------------------------------------------
# fit a model

# kris fit yr, hatchery, avg_weight, jday w/ separate tag_codes
# model <-
#   gam(
#     cbind(total_cwt_recovered, total_cwt_rel - total_cwt_recovered) ~
#       s(brood_year, bs="cs", k=length(min(dat$brood_year):max(dat$brood_year))) +
#       hatchery_group +
#       s(hatchery, bs="re") +
#       s(avg_weight, bs = "tp", k = 3) +
#       s(jday, bs = "tp") +
#       s(hatchery, brood_year, bs="fs",k=-1),
#     family = quasibinomial(link = "logit"),
#     data = dat,
#     method="REML"
#   )


########################## Join Release & Return Data for SAR ##########################
sar_df <-left_join(rel_agg, rec_join[,c("tag_code","number_cwt_estimated")], by = "tag_code") %>%
         mutate(avg_weight = round(avg_weight, digits = -1)) %>%
         aggregate(cbind(event_released,number_cwt_estimated)~species+brood_year+release_location_code+hatchery_location_code+
                                                              first_release_date_month+jday+avg_weight+release_stage,
                   data=., FUN=sum)

sar_df$sar <-sar_df$number_cwt_estimated/sar_df$event_released

# sar_df$release_location_code <-as.factor(sar_df$release_location_code)
# sar_df$first_release_date_month <-as.factor(sar_df$first_release_date_month)
# sar_df$avg_weight <-as.factor(sar_df$avg_weight)
sar_df <-filter(sar_df,release_stage!="no_stage")
sar_df$tags_lost <-sar_df$event_released - sar_df$number_cwt_estimated
write.csv(sar_df, paste0(proj_drive))

##### Univariate
### brood_year: 42.9%
model_by <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp"), family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_by)
### avg_weight: 14.6%
# model_wt <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(avg_weight, bs = "tp", k = 3), family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_wt)
### jday: 12%
# model_jd <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(jday, bs = "tp"), family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_jd)
### first_release_date_month: 9.21%
# model_rd <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(first_release_date_month, bs = "re"),family=quasibinomial(link="logit"),data=sar_df,method="REML")
# summary(model_rd)


##### Bivariate
### brood_year + avg_weight: 52.3%
model_bywt <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k = 3),
               family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywt)
### brood_year + jday: 46%
model_byjd <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(jday, bs = "cc"),
               family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_byjd)
### brood_year + first_release_date_month: 77.1%
model_byrd <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(first_release_date_month, bs = "re"),
               family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_byrd)
### avg_weight + jday: 16.5%
# model_wtjd <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(avg_weight, bs = "tp", k = 3) + s(jday, bs = "tp"),
#                family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_wtjd)
### avg_weight + first_release_date_month: 13.5%
# model_wtrd <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(avg_weight,bs="tp",k=3)+s(first_release_date_month,bs="re"),
#                family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_wtrd)
### jday + first_release_date_month: 30.6%
# model_jdrd <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(jday,bs="tp")+s(first_release_date_month,bs="re"),
#                family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_jdrd)


##### Multivariate
### brood_year + avg_weight + jday: 52.2%
model_bywtjd <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k = 3) + s(jday, bs = "cc"),
               family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtjd)
### brood_year + avg_weight + first_release_date_month: 52.6%
model_bywtrd <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k = 3) + s(first_release_date_month,bs="re"),
               family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtrd)
### brood_year + jday + first_release_date_month: 58%
model_byjdrd <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(jday, bs = "cc") + s(first_release_date_month,bs="re"),
               family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_byjdrd)
### jday + avg_weight + first_release_date_month: 37.4%
# model_jdwtrd <-gam(cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~ s(jday, bs = "tp") + s(avg_weight, bs = "tp", k = 3) + s(first_release_date_month,bs="re"),
#                family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_jdwtrd)


##### Multivariate w/ interaction
###
### brood_year + avg_weight + jday + jdayXbrood_year: 94.4%
model_bywtjd_int1 <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k = 3) + s(jday, bs = "cc") +
                       te(jday, brood_year, bs="fs"),
                    family = quasibinomial(link = "logit"), data = sar_df, method="REML")
summary(model_bywtjd_int1)
### brood_year + avg_weight + jday + avg_weightXbrood_year: 59.4%
model_bywtjd_int2 <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k = 3) + s(jday, bs = "cc") +
                       te(avg_weight, brood_year, bs="fs", k=3),
                    family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtjd_int2)
### brood_year + avg_weight + jday + first_release_date_monthXbrood_year: 94.7%
model_bywtjd_int3 <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k = 3) + s(jday, bs = "cc") +
                       te(first_release_date_month, brood_year, bs="re"),
                    family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtjd_int3)
###
### brood_year + avg_weight + first_release_date_month + jdayXbrood_year: 94.8%
model_bywtrd_int1 <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k=3)+s(first_release_date_month,bs="re") +
                       te(jday, brood_year, bs="fs"),
                    family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtrd_int1)
### brood_year + avg_weight + first_release_date_month + avg_weightXbrood_year: 55.3%
model_bywtrd_int2 <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k=3)+s(first_release_date_month,bs="re") +
                       te(avg_weight, brood_year, bs="fs",k=3),
                    family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtrd_int2)
### brood_year + avg_weight + first_release_date_month + first_release_date_monthXbrood_year: 52.6%
model_bywtrd_int3 <-gam(cbind(number_cwt_estimated, tags_lost) ~ s(brood_year, bs = "tp") + s(avg_weight, bs = "tp", k=3)+s(first_release_date_month,bs="re") +
                       te(first_release_date_month, brood_year, bs="re"),
                    family = quasibinomial(link = "logit"), data = sar_df, method="REML")
# summary(model_bywtrd_int3)
# model_bywtjd_int1
### gcv_ubre used instead of AIC
gcv_ubre <-data.frame(gcv_ubre = as.numeric(c(model_by$gcv.ubre,
           model_bywt$gcv.ubre, model_byjd$gcv.ubre, model_byrd$gcv.ubre,
           model_bywtjd$gcv.ubre, model_bywtrd$gcv.ubre, model_byjdrd$gcv.ubre,
           model_bywtjd_int1$gcv.ubre, model_bywtjd_int2$gcv.ubre, model_bywtjd_int3$gcv.ubre,
           model_bywtrd_int1$gcv.ubre, model_bywtrd_int2$gcv.ubre, model_bywtrd_int3$gcv.ubre)),
           trt = c("model_by",
           "model_bywt", "model_byjd", "model_byrd",
           "model_bywtjd", "model_bywtrd", "model_byjdrd",
           "model_bywtjd_int1", "model_bywtjd_int2", "model_bywtjd_int3",
           "model_bywtrd_int1", "model_bywtrd_int2", "model_bywtrd_int3"))
gcv_ubre = gcv_ubre[order(gcv_ubre$gcv_ubre),]
gcv_ubre[1,"trt"]
best.model <-model_bywtjd_int1

# model_bywtrd_int1 wins
summary(best.model)


# #####
# # model3: te(release_month) + s(avg_weight) + te(jday,brood_year) == 95% deviance
# model3 <-  gam(
#             cbind(number_cwt_estimated, event_released - number_cwt_estimated) ~
#               s(brood_year, bs="cs", k=-1) +
#               s(avg_weight, bs = "cs", k = 3) +
#               te(first_release_date_month, bs = "re") +
#               te(jday, brood_year, bs="fs",k=-1),
#             family = quasibinomial(link = "logit"),
#             data = sar_df,
#             method="REML"
#           )
# summary(model3)




# Make predictions without including "tag_code"
preds <-predict(best.model, newdata = sar_df, type = "response", se.fit = T) |>
        as_tibble() |>
        rename(prediction = fit, pred_se = se.fit) |>
        mutate(lci = qnorm(0.025, prediction, pred_se),
               uci = qnorm(0.975, prediction, pred_se),
               across(lci, ~ case_when(. < 0 ~ 0, .default = .)))

# add predictions
sar_pred <-sar_df |> bind_cols(preds)
```

## Plot Predicted SAR

Generate figures showing the results of the SAR calculations and the selected model fit.

```{r plot}
#create some plots
ggplot(sar_pred, aes(x=brood_year,y=sar,group=first_release_date_month,color=as.factor(first_release_date_month))) +
  geom_line(aes(x = brood_year, y = prediction,group=first_release_date_month)) +  # Line plot to show trends over brood years
  geom_ribbon(aes(ymin = lci,ymax = uci,group=first_release_date_month),alpha = 0.05,color = "grey") +
  geom_point(aes(group=first_release_date_month)) + # Add points for individual predictions
  facet_wrap(.~avg_weight, ncol=1)+
# ggplot(sar_pred, aes(x=brood_year,y=sar,group=avg_weight,color=avg_weight)) +
#   geom_line(aes(x = brood_year, y = prediction,group=avg_weight)) +  # Line plot to show trends over brood years
#   geom_ribbon(aes(ymin = lci,ymax = uci,group=avg_weight),alpha = 0.05,color = "grey") +
#   geom_point(aes(group=avg_weight)) + # Add points for individual predictions
#   facet_wrap(.~first_release_date_month,ncol=1)+
  # theme_minimal() +
  labs(x = "Brood Year", y = "Predicted SAR") +
  scale_x_continuous(breaks=seq(min(sar_pred$brood_year),max(sar_pred$brood_year),1))+
  # scale_y_log10(breaks = scales::breaks_pretty()) +
  # scale_y_continuous(trans = "log",
  #                    breaks = scales::breaks_pretty()) +
  theme(legend.position = "right")


# # add mean observed values by hatchery / brood year
# all_preds <-newdat |>
#   select(hatchery, hatchery_group, brood_year) |>
#   left_join(dat |> group_by(hatchery, brood_year) |>
#   summarize(across(c(avg_weight, jday, relative_sar), ~ mean(., na.rm = T)), .groups = "drop"), 
#             by = join_by(hatchery, brood_year)) |> 
#   filter(!is.na(relative_sar))
# 
# all_preds <-all_preds |> 
#   bind_cols(predict(model, newdata = all_preds, type = "response", se.fit = T) |>
#   as_tibble() |>
#   rename(prediction = fit, pred_se = se.fit) |>
#   mutate(lci = qnorm(0.025, prediction, pred_se),
#          uci = qnorm(0.975, prediction, pred_se),
#          across(lci, ~ case_when(. < 0 ~ 0, .default = .))))

sar_pred |> 
  ggplot(aes(x = sar, y = prediction)) + 
  geom_abline(linetype = 2) +
  geom_point(size = 3) +
  # scale_x_log10(breaks = scales::breaks_pretty()) +
  # scale_y_log10(breaks = scales::breaks_pretty()) +
  labs(x = "Observed SAR", y = "Predicted SAR")


dat <- ggpredict(best.model, terms = c("jday", "avg_weight"))
plot(dat, facet = TRUE)

dat <- ggpredict(best.model, terms = c("brood_year", "avg_weight"))
plot(dat, facet = TRUE)

dat <- ggpredict(best.model, terms = c("jday", "brood_year","avg_weight"))
plot(dat, facet = TRUE)

dat <- ggpredict(best.model, terms = c("brood_year", "jday"))
plot(dat, facet = TRUE)


```
